Time Stretching of the GeV Emission of GRBs: Fermi LAT data vs geometrical model

### 2014-03-07 Paper/gammaRays.tex
change TIME_EXTENSION_FACTOR from 1.3 to 2.0
KS-test drafted

\begin{figure}[modelOverview]
        \centering
        \includegraphics[width=0.7\textwidth]{modelOverview}
        \caption{
        	Model Overview. Here the red and the blue cones represent the volumes through which low and high energy plasma propagates. In case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so, due to the relativistic beaming effect, the most of the observable low energy photons will travel along the straight line from the central engine. Also, the observer's off-axis angle is larger that the opening angle of the high energy jet, so the high energy radiation will still originate near the center of the jet (because it is the only place where there are high energy radiators). The observation time of a photon is a sum of two things: the time interval spent in plasma as a radiator (which approximately equals to the distance from the central engine to the point of emission); and the time interval from emission to detection (which is the distance from the point of emission to the observer's location). Given a position of the shock front, this sum is larger for high energy photons. Because of that, high energy emission will be observed later throughout the burst duration, therefore the high energy light curve will be stretched.
        }
        \label{fig:modelOverview}
\end{figure}

\begin{figure}[sampleLightCurve]
	\centering
	\includegraphics[width=0.49\textwidth]{sampleLightCurve}
	\caption{
		High and low energy light curves produced by the geometrical model. Parameter values are the same as discussed in the section \ref{sec:parameters}. Redshift $z = 1$. Observer's off-axis angle $\chi = 0.01$.
	}
	\label{fig:sampleLightCurve}
\end{figure}

@@

\begin{table}[bursts]
	\centering
	\small
	\begin{tabular}{ l | S[table-format=9.3] | S[table-format=3.3] | S[table-format=-2.3] | S[table-format=1.5] | S[table-format=2.1] | S[table-format=3.1] }
		\multicolumn{1}{ c |}{\texttt{name}} & $\mathrm{\texttt{time}}$ & \texttt{location.ra} & \texttt{location.dec} & \texttt{location.err} & \texttt{startOffset} & \texttt{endOffset} \\
		\multicolumn{1}{ c |}{GRB} & $\mathrm{GBM\ Trigger\ Time}$ & $\mathrm{R.A.}$ & $\mathrm{Dec.}$ & $\mathrm{Location\ Error}$ & $T_{05}$ & $T_{95}$ \\
		\multicolumn{1}{ c |}{Name}& $\mathrm{MET,\ sec}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{deg}$ & $\mathrm{sec}$ & $\mathrm{sec}$ \\
		\hline
		080825C	&	241366429.105	&	233.9  	&	 -4.5  	&	0.75   	&	 3.2	&	 29.4	\\
		080916C	&	243216766.614	&	119.85 	&	-56.64 	&	0.0001 	&	 5.0	&	209.8	\\
		081006 	&	244996175.173	&	136.32 	&	-62.05 	&	0.52   	&	 0.7	&	115.0	\\
		081024B	&	246576161.864	&	322.95 	&	 21.2  	&	0.22   	&	 0.1	&	191.0	\\
		090217 	&	256539404.560	&	204.83 	&	 -8.42 	&	0.35   	&	 6.2	&	 68.0	\\
		090323 	&	259459364.630	&	190.71 	&	 17.053	&	0.0001 	&	15.9	&	293.9	\\
		090328 	&	259925808.510	&	 90.67 	&	-41.715	&	0.0002 	&	18.8	&	652.9	\\
		090510 	&	263607781.971	&	333.55 	&	-26.583	&	0.0004 	&	 0.6	&	 45.6	\\
		090626 	&	267683530.880	&	170.03 	&	-33.49 	&	0.22   	&	52.2	&	299.9	\\
		090902B	&	273582310.313	&	264.94 	&	 27.324	&	0.001  	&	 7.7	&	825.0	\\
		090926A	&	275631628.990	&	353.4  	&	-66.32 	&	0.01   	&	 5.5	&	225.0	\\
		091003 	&	276237347.585	&	251.52 	&	 36.625	&	0.0005 	&	 3.9	&	452.6	\\
		091031 	&	278683230.850	&	 71.49 	&	-57.65 	&	0.23   	&	 3.1	&	206.2	\\
		100116A	&	285370262.240	&	305.01 	&	 14.43 	&	0.17   	&	 3.0	&	141.0	\\
		100414A	&	292904423.990	&	192.11 	&	  8.693	&	0.0005 	&	17.4	&	288.6	\\
		110120A	&	317231981.230	&	 61.5  	&	-12.0  	&	0.36   	&	 0.5	&	112.8	\\
		110428A	&	325675112.410	&	  5.59 	&	 64.849	&	0.00001	&	10.7	&	407.6	\\
		110721A	&	332916465.760	&	333.2  	&	-38.5  	&	0.20   	&	 0.1	&	239.0	\\
		110731A	&	333803371.954	&	280.504	&	-28.537	&	0.0001 	&	 3.0	&	 24.1
	\end{tabular}
	\caption{Burst data used in our study. The first row contains the corresponding \texttt{GRBurst} class variable names. The data is taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}. \url{https://github.com/maxitg/GammaRays/blob/master/GRObservations/bursts}}
	\label{tab:bursts}
\end{table}

@@

Knowing the exposures, we can now use the method introduced in the appendix of \cite{Rubtsov:2011qq} to estimate the linear component of background from the data observed before the burst. We compute two background estimates: for low and high energy bands.

\subsection{KS-test}
\subsection{Results}

>>>>

\begin{figure}
        \centering
        \includegraphics[width=0.7\textwidth]{modelOverview}
        \caption{
        	Model Overview. Here the red and the blue cones represent the volumes through which low and high energy plasma propagates. In case depicted the observer's off-axis angle is smaller than the opening angle of the low energy jet, so, due to the relativistic beaming effect, the most of the observable low energy photons will travel along the straight line from the central engine. Also, the observer's off-axis angle is larger that the opening angle of the high energy jet, so the high energy radiation will still originate near the center of the jet (because it is the only place where there are high energy radiators). The observation time of a photon is a sum of two things: the time interval spent in plasma as a radiator (which approximately equals to the distance from the central engine to the point of emission); and the time interval from emission to detection (which is the distance from the point of emission to the observer's location). Given a position of the shock front, this sum is larger for high energy photons. Because of that, high energy emission will be observed later throughout the burst duration, therefore the high energy light curve will be stretched.
        }
        \label{fig:modelOverview}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.49\textwidth]{sampleLightCurve}
	\caption{
		High and low energy light curves produced by the geometrical model. Parameter values are the same as discussed in the section \ref{sec:parameters}. Redshift $z = 1$. Observer's off-axis angle $\chi = 0.01$.
	}
	\label{fig:sampleLightCurve}
\end{figure}

@@

\begin{table}
	\centering
	\small
	\begin{tabular}{ l | S[table-format=9.3] | S[table-format=3.3] | S[table-format=-2.3] | S[table-format=1.5] | S[table-format=2.1] | S[table-format=3.1] }
		\multicolumn{1}{ c |}{\texttt{name}} & $\mathrm{\texttt{time}}$ & \texttt{location.ra} & \texttt{location.dec} & \texttt{location.err} & \texttt{startOffset} & \texttt{endOffset} \\
		\multicolumn{1}{ c |}{GRB} & $\mathrm{GBM\ Trigger\ Time}$ & $\mathrm{R.A.}$ & $\mathrm{Dec.}$ & $\mathrm{Location\ Error}$ & $T_{05}$ & $T_{95}$ \\
		\multicolumn{1}{ c |}{Name}& $\mathrm{MET,\ sec}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{J2000,\ deg}$ & $\mathrm{deg}$ & $\mathrm{sec}$ & $\mathrm{sec}$ \\
		\hline
		080825C	&	241366429.105	&	233.9  	&	 -4.5  	&	0.75   	&	 3.2	&	 29.4	\\
		080916C	&	243216766.614	&	119.85 	&	-56.64 	&	0.0001 	&	 5.0	&	209.8	\\
		081006 	&	244996175.173	&	136.32 	&	-62.05 	&	0.52   	&	 0.7	&	115.0	\\
		081024B	&	246576161.864	&	322.95 	&	 21.2  	&	0.22   	&	 0.1	&	191.0	\\
		090217 	&	256539404.560	&	204.83 	&	 -8.42 	&	0.35   	&	 6.2	&	 68.0	\\
		090323 	&	259459364.630	&	190.71 	&	 17.053	&	0.0001 	&	15.9	&	293.9	\\
		090328 	&	259925808.510	&	 90.67 	&	-41.715	&	0.0002 	&	18.8	&	652.9	\\
		090510 	&	263607781.971	&	333.55 	&	-26.583	&	0.0004 	&	 0.6	&	 45.6	\\
		090626 	&	267683530.880	&	170.03 	&	-33.49 	&	0.22   	&	52.2	&	299.9	\\
		090902B	&	273582310.313	&	264.94 	&	 27.324	&	0.001  	&	 7.7	&	825.0	\\
		090926A	&	275631628.990	&	353.4  	&	-66.32 	&	0.01   	&	 5.5	&	225.0	\\
		091003 	&	276237347.585	&	251.52 	&	 36.625	&	0.0005 	&	 3.9	&	452.6	\\
		091031 	&	278683230.850	&	 71.49 	&	-57.65 	&	0.23   	&	 3.1	&	206.2	\\
		100116A	&	285370262.240	&	305.01 	&	 14.43 	&	0.17   	&	 3.0	&	141.0	\\
		100414A	&	292904423.990	&	192.11 	&	  8.693	&	0.0005 	&	17.4	&	288.6	\\
		110120A	&	317231981.230	&	 61.5  	&	-12.0  	&	0.36   	&	 0.5	&	112.8	\\
		110428A	&	325675112.410	&	  5.59 	&	 64.849	&	0.00001	&	10.7	&	407.6	\\
		110721A	&	332916465.760	&	333.2  	&	-38.5  	&	0.20   	&	 0.1	&	239.0	\\
		110731A	&	333803371.954	&	280.504	&	-28.537	&	0.0001 	&	 3.0	&	 24.1
	\end{tabular}
	\caption{Burst data used in our study. The first row contains the corresponding \texttt{GRBurst} class variable names. The data is taken from the tables 2 and 4 of \cite{Ackermann:2013zfa}. \url{https://github.com/maxitg/GammaRays/blob/master/GRObservations/bursts}}
	\label{tab:bursts}
\end{table}

@@

Knowing the exposures, we can now use the method introduced in the appendix of \cite{Rubtsov:2011qq} to estimate the linear component of background from the data observed before the burst. We compute two background estimates: for low and high energy bands. Because the background of transient class photons may depend on the spacecraft position, we have to filter them as well.

\subsection{KS-test}

We take care of remaining background by using the following functions with the Kolmogorov-Smirnov test:
\begin{equation}
	\Phi_i\left(t\right) = \frac{p_i\left(T_1, t\right) - b_i \frac{t-T_1}{T_2-T_1}}{p_i\left(T_1, T_2\right) - b_i}
\end{equation}
Here $\left(T_1, T_2\right)$ is the time range of observations, $p_i\left(t_1, t_2\right)$ is the number of photons observed in the time range $t_1$ to $t_2$ in the $i$'s energy band ($i=\text{low},\text{high}$ are for the low and high energy bands respectfully), and $b_i$ is the estimated number of background photons for the time range $T_1$ to $T_2$ in the $i$'s energy band.

The numbers of degrees of freedom (used as an input for the KS-test) are $p_i\left(T_1, T_2\right) - b_i$.

While we did not prove it, we assume that the Kolmogorov-Smirnov statistics computed for these functions with provided number of degrees of freedom has a similar distribution to that of the KS-statistics for ordinary CDFs (that is CDFs with no background).

Finally, we compute KS-probabilities for pairs $\Phi_\text{low}\left(t\right), \Phi_\text{high}\left(\kappa t\right)$ to obtain the allowed ranges for stretching factors $\kappa$. We evaluate stretching factors in the range $\texttt{STRETCHING\_MIN} = 0.1$ to $\texttt{STRETCHING\_MAX} = 10.$ with a logarithmic step of $\texttt{STRETCHING\_STEP} = 1.001$.

As a result of such calculation we can constraint the range of allowed stretching factors to values for which the KS-probabilities are not too small, in other words, given the sigma-significance value, compute the range outside of which stretching factors are elliminated by observations.

You can see the results of such computation in the following section.
